---
title: "Homework 3"
date: January 26, 2022
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

- This homework is due Wednesday, February 2 at 3pm EST.
- Submit via GitHub.  Remember to commit and push online so I can see it.
- Please format your homework solutions using R Markdown.  You are welcome to simply add your answers below each question.
  - If the question requires a figure, make sure you have informative title, axis labels, and legend if needed.
  - Note: When I've given the framework of an answer's code, I've included the option `eval=FALSE` in the R chunk.  When you start filling in your answer, you'll need to switch this to `eval=TRUE`.
- Turn in both the .rmd file and the knitted .pdf or .html file.
  - Knitting the .rmd file to a .pdf or .html file should help ensure your code runs without errors, but double check the output is what you expected.

## Question 1

- Sampling variability
- Sampling distribution
- Standard deviation
- Standard error

For each of these terms, in your own words:

1. explain what it means in terms of $\hat{ATE}$.
2. explain how it relates to the other terms listed.

(Don't be afraid of repeating yourself!)


## Question 2

Read the Exercise 8 prompt from GG Ch. 3.  Although a natural experiment, explain in your own words how this data was generated, in effect, using a block randomized design.  In other words, why should we analyze this data using what we know about estimating $\hat{ATE}$ and $\hat{SE}(\hat{ATE})$ under block randomization?



## Question 3

Complete Exercise 8 from GG Ch. 3, parts a, b, c, and e.

Note: do these questions "by hand," meaning, calculate the quantities on your own in code without using pre-programmed statistical routines like `lm()`.  (You can still use commands like `mean()`!)

In this code chunk, I'm just creating the dataset shown in the book for you.

```{r}
term_length_tx <- c(rep(0, 16),
                    rep(1, 15))
term_length_ak <- c(rep(0, 16),
                    rep(1, 18), 0)
bills_tx <- c(18,29,41,53,60,67,75,79,79,
              88,93,101,103,106,107,131,
              29,37,42,45,45,54,54,58,61,
              64,69,73,75,92,104)
bills_ak <- c(11,15,23,24,25,26,28,31,33,
              34,35,35,36,38,52,59,9,
              10,14,15,15,17,18,19,19,
              20,21,23,23,24,28,30,32,34,17)
df <- data.frame(state = c(rep("TX", 31),
                           rep("AK", 35)),
                 term_length = c(term_length_tx, term_length_ak),
                 bills = c(bills_tx, bills_ak))
```


### 3a



### 3b



### 3c



### 3e



## Question 4

Now, answer Exercise 8 part d.

In addition to explaining what the question asks you to explain, calculate $\hat{ATE}$ when pooling and compare the estimated ATE to when you didn't pool in 3c.

Also, calculate $\hat{SE}(\hat{ATE})$ when pooling and compare the estimated ATE to when you didn't pool in 3e.

What are the implications if the researcher incorrectly analyzes this data?


# Checking your answers

Confirm your correct results from Question 3 and your incorrect results from Question 4 using software commands available in R.  I've provided the commands for you.  If what you calculated in Question 3 and Question 4 matches up, then you know what's happening under the hood in these functions!

```{r}
# with blocking
estimatr::difference_in_means(bills ~ term_length,
                              blocks = state,
                              data = df)
# without analyzing within-block first, then aggregating
estimatr::difference_in_means(bills ~ term_length,
                              data = df)
```

